---
title: "logistic_regression_prediction"
output: html_document
---

# top results from top diseases univariate regressions

```{r}

log_top_1=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/top_train1_mr.rds")
perf_top_1=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/top_test_1_mr.rds")

```

```{r}
summary(log_top_1)
```

    FALSE TRUE
  0  4553  469
  1  3737 1285
[1] 0.5868597


```{r}
plot(perf_top_1,colorize=TRUE)
```
```{r}
log_top_5=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/top_train5_mr.rds")
perf_top_5=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/top_test_5_mr.rds")
```

```{r}
summary(log_top_5)
```

    FALSE TRUE
  0  4745  277
  1  4251  771
[1] 0.5523043



```{r}
plot(perf_top_5,colorize=TRUE)
```
# on all comorbidities

# 1 year prior
  
    FALSE TRUE
  0  4348  674
  1  3841 1181
[1] 0.5504779

```{r}

log_1=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/train1_mr.rds")
perf_1=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/test_1_mr.rds")

```


```{r}
plot(perf_1,colorize=TRUE)
```

# 5 years

```{r}

log_5=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/train5_mr.rds")
perf_5=readRDS( "/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/test_5_mr.rds")

```


    FALSE TRUE
  0  1018 4004
  1  1141 3881
[1] 0.4877539

```{r}
plot(perf_5,colorize=TRUE)
```


# Using sun exposures only

```{r}

train=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/train_sun_exp.rds")

colnames(train)

train$sex=as.factor(train$sex)
train$childhood_sunburn=as.numeric(train$childhood_sunburn)
train$family_history_cancer=as.factor(train$family_history_cancer)
train$ease_of_skin_tanning=as.factor(train$ease_of_skin_tanning)
train$time_spent_outdoors_summer=as.numeric(train$time_spent_outdoors_summer)
train$skin_colour=as.factor(train$skin_colour)

test=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/test_sun_exp.rds")

test$sex=as.factor(test$sex)
test$childhood_sunburn=as.numeric(test$childhood_sunburn)
test$family_history_cancer=as.factor(test$family_history_cancer)
test$ease_of_skin_tanning=as.factor(test$ease_of_skin_tanning)
test$time_spent_outdoors_summer=as.numeric(test$time_spent_outdoors_summer)
test$skin_colour=as.factor(test$skin_colour)



library(gplots)
library(caTools)
library(ROCR)

set.seed(31)

log_reg_pred = glm(status ~ .,  data = train, family=binomial)
summary(log_reg_pred)

predictTest = predict(log_reg_pred, type="response", newdata=test)
table(test$status, predictTest > 0.5)

ROCRpred = prediction(predictTest, test$status)
exp_perf_auc=as.numeric(performance(ROCRpred, "auc")@y.values)
exp_perf_auc

exp_perf <- performance(ROCRpred,"tpr","fpr")


```
```{r}
plot(exp_perf,colorize=TRUE)
```

# all covariates

```{r}
train=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/train_all_covariates.rds")

train=train[,c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")]


train$sex=as.factor(train$sex)
train$age=as.numeric(train$age)
train$bmi=as.numeric(train$bmi)
train$health_rating=as.factor(train$health_rating)
train$height=as.numeric(train$height)
train$childhood_sunburn=as.numeric(train$childhood_sunburn)
train$smoking_status=as.factor(train$smoking_status)
train$alcohol_intake_frequency=as.factor(train$alcohol_intake_frequency)
train$skin_colour=as.factor(train$skin_colour)
train$ease_of_skin_tanning=as.factor(train$ease_of_skin_tanning)
train$time_spent_outdoors_summer=as.numeric(train$time_spent_outdoors_summer)
train$vegetable_intake=as.numeric(train$vegetable_intake)
train$water_intake=as.numeric(train$water_intake)
train$family_history_cancer=as.factor(train$family_history_cancer)


#covariates_interest= c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")

test=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/test_all_covariates.rds")

test=test[,c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")]



test$sex=as.factor(test$sex)
test$age=as.numeric(test$age)
test$bmi=as.numeric(test$bmi)
test$health_rating=as.factor(test$health_rating)
test$height=as.numeric(test$height)
test$childhood_sunburn=as.numeric(test$childhood_sunburn)
test$smoking_status=as.factor(test$smoking_status)
test$alcohol_intake_frequency=as.factor(test$alcohol_intake_frequency)
test$skin_colour=as.factor(test$skin_colour)
test$ease_of_skin_tanning=as.factor(test$ease_of_skin_tanning)
test$time_spent_outdoors_summer=as.numeric(test$time_spent_outdoors_summer)
test$vegetable_intake=as.numeric(test$vegetable_intake)
test$water_intake=as.numeric(test$water_intake)
test$family_history_cancer=as.factor(test$family_history_cancer)



library(gplots)
library(caTools)
library(ROCR)

set.seed(31)

log_reg_pred = glm(status ~ .,  data = train, family=binomial)
summary(log_reg_pred)

predictTest = predict(log_reg_pred, type="response", newdata=test)
table(test$status, predictTest > 0.5)

ROCRpred = prediction(predictTest, test$status)
perf_auc_covariates=as.numeric(performance(ROCRpred, "auc")@y.values)
perf_auc_covariates

perf_covariates <- performance(ROCRpred,"tpr","fpr")

```

```{r}
plot(perf,colorize=TRUE)
```




# top comorbidities + covariates + biom 1 year prior

```{r}
train_com<-readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/train1.rds")
test_com<-readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/test1.rds")

ICD10_1<-readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/ICD10_1.rds")
new_train<- train_com[ ,which(colnames(train_com) %in% ICD10_1)]
new_train<-cbind(train_com['status'], new_train)

new_test<- test_com[ ,which(colnames(test_com) %in% ICD10_1)]
new_test<-cbind(test_com['status'], new_test)


train=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/train_all_covariates.rds")

train=train[,c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")]


train$sex=as.factor(train$sex)
train$age=as.numeric(train$age)
train$bmi=as.numeric(train$bmi)
train$health_rating=as.factor(train$health_rating)
train$height=as.numeric(train$height)
train$childhood_sunburn=as.numeric(train$childhood_sunburn)
train$smoking_status=as.factor(train$smoking_status)
train$alcohol_intake_frequency=as.factor(train$alcohol_intake_frequency)
train$skin_colour=as.factor(train$skin_colour)
train$ease_of_skin_tanning=as.factor(train$ease_of_skin_tanning)
train$time_spent_outdoors_summer=as.numeric(train$time_spent_outdoors_summer)
train$vegetable_intake=as.numeric(train$vegetable_intake)
train$water_intake=as.numeric(train$water_intake)
train$family_history_cancer=as.factor(train$family_history_cancer)


covariates_interest= c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")

test=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/test_all_covariates.rds")

test=test[,c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")]


test$sex=as.factor(test$sex)
test$age=as.numeric(test$age)
test$bmi=as.numeric(test$bmi)
test$health_rating=as.factor(test$health_rating)
test$height=as.numeric(test$height)
test$childhood_sunburn=as.numeric(test$childhood_sunburn)
test$smoking_status=as.factor(test$smoking_status)
test$alcohol_intake_frequency=as.factor(test$alcohol_intake_frequency)
test$skin_colour=as.factor(test$skin_colour)
test$ease_of_skin_tanning=as.factor(test$ease_of_skin_tanning)
test$time_spent_outdoors_summer=as.numeric(test$time_spent_outdoors_summer)
test$vegetable_intake=as.numeric(test$vegetable_intake)
test$water_intake=as.numeric(test$water_intake)
test$family_history_cancer=as.factor(test$family_history_cancer)

rownames(biomarker_final)=biomarker_final$eid

new_test=new_test[which( rownames(new_test) %in% rownames(biomarker_final)),]
new_train=new_train[which( rownames(new_train)%in% rownames(biomarker_final)) ,]

test=test[which( rownames(test) %in% rownames(new_test)),]
train=train[which( rownames(train)%in% rownames(new_train)) ,]

train_biom= biomarker_final[which(rownames(biomarker_final) %in% rownames(new_train)),]
train_biom=train_biom[,2:29]

test_biom= biomarker_final[which( rownames(biomarker_final) %in%  rownames(new_test)),]
test_biom=test_biom[,2:29]

test=cbind(new_test,test, test_biom)
train=cbind(new_train,train,train_biom)

```

```{r}
set.seed(31)

log_reg_pred = glm(status ~ .,  data = train, family=binomial)
summary(log_reg_pred)

predictTest = predict(log_reg_pred, type="response", newdata=test)
table(test$status, predictTest > 0.5)

ROCRpred = prediction(predictTest, test$status)
top_com_cov_auc_1=as.numeric(performance(ROCRpred, "auc")@y.values)
top_com_cov_auc_1

perf_com_cov_1 <- performance(ROCRpred,"tpr","fpr")


```

```{r}
plot(perf,colorize=TRUE)
```

Covariates+ top diseases 5 years


```{r}
train_com<-readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/train5.rds")
test_com<-readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/test5.rds")

ICD10_5<-readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Logistic_reg/ICD10_5.rds")
new_train<- train_com[ ,which(colnames(train_com) %in% ICD10_5)]
new_train<-cbind(train_com['status'], new_train)

new_test<- test_com[ ,which(colnames(test_com) %in% ICD10_5)]
new_test<-cbind(test_com['status'], new_test)


train=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/train_all_covariates.rds")

train=train[,c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")]


train$sex=as.factor(train$sex)
train$age=as.numeric(train$age)
train$bmi=as.numeric(train$bmi)
train$health_rating=as.factor(train$health_rating)
train$height=as.numeric(train$height)
train$childhood_sunburn=as.numeric(train$childhood_sunburn)
train$smoking_status=as.factor(train$smoking_status)
train$alcohol_intake_frequency=as.factor(train$alcohol_intake_frequency)
train$skin_colour=as.factor(train$skin_colour)
train$ease_of_skin_tanning=as.factor(train$ease_of_skin_tanning)
train$time_spent_outdoors_summer=as.numeric(train$time_spent_outdoors_summer)
train$vegetable_intake=as.numeric(train$vegetable_intake)
train$water_intake=as.numeric(train$water_intake)
train$family_history_cancer=as.factor(train$family_history_cancer)


#covariates_interest= c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")

test=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/Model_dataset/test_all_covariates.rds")

test=test[,c("status","age" ,"bmi", "sex", "health_rating" , "height" ,"childhood_sunburn" , "smoking_status" ,"alcohol_intake_frequency",   "skin_colour" ,"ease_of_skin_tanning"    ,   "time_spent_outdoors_summer", "vegetable_intake",   "water_intake",   "family_history_cancer")]


test$sex=as.factor(test$sex)
test$age=as.numeric(test$age)
test$bmi=as.numeric(test$bmi)
test$health_rating=as.factor(test$health_rating)
test$height=as.numeric(test$height)
test$childhood_sunburn=as.numeric(test$childhood_sunburn)
test$smoking_status=as.factor(test$smoking_status)
test$alcohol_intake_frequency=as.factor(test$alcohol_intake_frequency)
test$skin_colour=as.factor(test$skin_colour)
test$ease_of_skin_tanning=as.factor(test$ease_of_skin_tanning)
test$time_spent_outdoors_summer=as.numeric(test$time_spent_outdoors_summer)
test$vegetable_intake=as.numeric(test$vegetable_intake)
test$water_intake=as.numeric(test$water_intake)
test$family_history_cancer=as.factor(test$family_history_cancer)


test=cbind(new_test,test)
train=cbind(new_train,train)

```

```{r}
set.seed(31)

log_reg_pred = glm(status ~ .,  data = train, family=binomial)
summary(log_reg_pred)

predictTest = predict(log_reg_pred, type="response", newdata=test)
table(test$status, predictTest > 0.5)

ROCRpred = prediction(predictTest, test$status)
top_com_cov_auc_5=as.numeric(performance(ROCRpred, "auc")@y.values)

perf_com_cov_5 <- performance(ROCRpred,"tpr","fpr")

```

```{r}
plot(perf,colorize=TRUE)
```


#Comparative roc curves

# 1 year 

- top comorbidities :perf_top_1, auc 0.5868597
- all comorbidities, perf_1, auc= 0.5504779
perf_covariates, perf_auc_covariates
perf_com_cov_1, top_com_cov_auc_1

```{r}
perf_top_1_auc= 0.587
perf_1_auc=0.550
perf_top_5_auc=0.552
perf_5_auc=0.488
```


```{r}
# Plot ROC curves
jpeg("pred_1_year.jpg", width = 800, height = 500)

plot(perf_top_1, col = "skyblue", main='Multivariate logistic regression (time= 1 year and above)')
plot(perf_1, add = TRUE, col = "darkgreen")
plot(perf_covariates, add = TRUE, col = "tomato")
plot(perf_com_cov_1, add = TRUE, col = "orange")
legend("bottomright", legend = c(paste("top comorbidities, AUC = ",
round(perf_top_1_auc, digits = 3)), paste("all comorbidities, AUC = ",
round(perf_1_auc, digits = 3)), paste("covariates only, AUC = ",
round(perf_auc_covariates, digits = 3)),  paste("top comorbidities + covariates, AUC = ",
round(top_com_cov_auc_1, digits = 3))), col = c("skyblue",
"darkgreen", "tomato", 'orange'), lty = 1)

dev.off()
```
```{r}
perf_1=perf_top_1
```

# 5 years

perf_top_5, auc=0.5523043
perf 5 , auc= 0.4877539
perf_covariates, perf_auc_covariates
perf_com_cov_5 , top_com_cov_auc_5

```{r}

jpeg("pred_5_year.jpg", width = 800, height = 500)
plot(perf_top_5, col = "skyblue",  main='Multivariate logistic regression (time= 5 years and above)')
plot(perf_5, add = TRUE, col = "darkgreen")
plot(perf_covariates, add = TRUE, col = "tomato")
plot(perf_com_cov_5, add = TRUE, col = "orange")

legend("bottomright", legend = c(paste("top comorbidities, AUC = ",
round(perf_top_5_auc, digits = 3)), paste("all comorbidities, AUC = ",
round(perf_5_auc, digits = 3)), paste("covariates only, AUC = ",
round(perf_auc_covariates, digits = 3)),  paste("top comorbidities + covariates, AUC = ",
round(top_com_cov_auc_5, digits = 3))), col = c("skyblue",
"darkgreen", "tomato", 'orange'), lty = 1)
dev.off()
```

