---
title: "Logistic_regression_analysis"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(ggplot2)
library(tidyr)
library(ukbtools)
```


## Results from uunivariate logistic regression analysis

Load results for different analyses

```{r}
results10<- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/results10_new2.rds")
results5<- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/results5_new2.rds")
results3<- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/results3_new2.rds")
results1<- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/results1_new2.rds")
results0<- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/results0_new2.rds")
results_coc<- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/results_coc_new2.rds")
```

For annotation 
```{r}
comorbidities_categories<- read.csv("/rds/general/project/hda_students_data/live/Group7/General/Final/comorbidities_categories.csv",sep=',')
```

# 10 years prior to dignosis (10+)
```{r}
results10$code=rownames(results10)
results10= merge(results10, comorbidities_categories, by='code')
results10<-results10%>%
  drop_na()
results10$ICD10=substr(results10$code, 1, 1)

```

```{r}
results10$logp=-log10(results10$pval)
bonf10=-log10(0.05/nrow(results10))
results10$significant=results10$logp>bonf10
significant_ICD_10=results10$code[which(results10$significant==TRUE)]
significant_pval=results10$pval[which(results10$significant==TRUE)]
meaning10=t(data.frame(data.frame(t(sapply(as.character(significant_ICD_10), FUN=ukb_icd_code_meaning)))$meaning))
ICD10_10=as.character(meaning10[,1])
final10=data.frame(ICD10_10,significant_pval)
final10 <- final10[order(significant_pval),]
final10 

saveRDS(final10 , "/rds/general/project/hda_students_data/live/Group7/General/Final/table_10.rds")
```
```{r}
dim(final10)
```

#12 diseases above significance level

```{r}
ten_year<-ggplot(
  results10, aes(as.numeric(row.names(results10)), y=-log10(pval), color=ICD10))+
  geom_point()+
  geom_hline(yintercept= -log10(0.05/nrow(results10)), linetype="dashed", color = "red")+
  geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(results10)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+
  ggtitle("Time= 10 years prior to diagnosis of skin cancer") +
  xlab("ICD10 codes")+
  ylim(0,60)+
  theme(axis.text.x = element_blank())+
  theme(aspect.ratio = 1)+
   theme(legend.box = 'horizontal')
#+theme(legend.position = "none")

ten_year
```

<!-- ```{r} -->
<!-- legend_description<-ggplot( -->
<!--   ten_bis, aes(as.numeric(row.names(ten_bis)), y=-log10(pval), color=Title))+ -->
<!--   geom_point()+ -->
<!--   geom_hline(yintercept= -log10(0.05/nrow(ten_bis)), linetype="dashed", color = "red")+ -->
<!--   geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(ten_bis)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+ -->
<!--   #ggtitle("Time= 10 years prior to diagnosis of skin cancer") + -->
<!--   xlab("ICD10 codes")+ -->
<!--   ylim(0,60)+ -->
<!--   theme(axis.text.x = element_blank())+ -->
<!--   theme(aspect.ratio = 1)+ -->
<!--   theme(legend.box = 'vertical')+ -->
<!--   theme_void()+ -->
<!--   theme(legend.key.size = unit(1, 'cm'), #change legend key size -->
<!--         legend.key.height = unit(1, 'cm'), #change legend key height -->
<!--         legend.key.width = unit(1.5, 'cm'), #change legend key width -->
<!--         legend.title = element_text(size=5), #change legend title font size -->
<!--         legend.text = element_text(size=8)) #change legend text font size -->
<!-- #+theme(legend.position = "none") -->

<!-- legend_description -->
<!-- ``` -->


```{r}
ggsave(filename="tenyear.pdf", 
       plot = ten_year, 
       device = cairo_pdf, 
       width = 210, 
       height = 297, 
       units = "mm")
```

```{r}
chap=read.csv('/rds/general/project/hda_students_data/live/Group7/General/Eleonore/chapters5.csv', sep=';')
chap
```


# 5 years prior to dignosis and above (5+)
```{r}
results5$code=rownames(results5)
results5= merge(results5, comorbidities_categories, by='code')
results5<-results5%>%
  drop_na()
results5$ICD10=substr(results5$code, 1, 1)
```

```{r}
results5$logp=-log10(results5$pval)
bonf5=-log10(0.05/nrow(results5))
results5$significant=results5$logp>bonf5
significant_ICD_5=results5$code[which(results5$significant==TRUE)]
significant_pval=results5$pval[which(results5$significant==TRUE)]
meaning5=t(data.frame(data.frame(t(sapply(as.character(significant_ICD_5), FUN=ukb_icd_code_meaning)))$meaning))
ICD10_5=as.character(meaning5[,1])
final5=data.frame(ICD10_5,significant_pval)
final5 <- final5[order(significant_pval),]
final5 

saveRDS(final5 , "/rds/general/project/hda_students_data/live/Group7/General/Final/table_5.rds")
```
```{r}
dim(final5)
```
#30 diseases above significance level


```{r}
five_year<-ggplot(
  results5, aes(as.numeric(row.names(results5)), y=-log10(pval), color=ICD10))+
  geom_point()+
  geom_hline(yintercept= -log10(0.05/nrow(results5)), linetype="dashed", color = "red")+
  geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(results5)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+
  ggtitle("Time= 5 years prior to diagnosis of skin cancer") +
  xlab("ICD10 codes")+
  ylim(0,60)+
  theme(axis.text.x = element_blank())+
  theme(aspect.ratio = 1)+
   theme(legend.box = 'horizontal')
#+theme(legend.position = "none")

five_year
```
```{r}
ggsave(filename="fiveyear.pdf", 
       plot = five_year, 
       device = cairo_pdf, 
       width = 210, 
       height = 297, 
       units = "mm")
```


# 3 years prior to dignosis and above (3+)

```{r}
results3$code=rownames(results3)
results3= merge(results3, comorbidities_categories, by='code')
results3<-results3%>%
  drop_na()
results3$ICD10=substr(results3$code, 1, 1)
```

```{r}
results3$logp=-log10(results3$pval)
bonf3=-log10(0.05/nrow(results3))
results3$significant=results3$logp>bonf3
significant_ICD_3=results3$code[which(results3$significant==TRUE)]
significant_pval=results3$pval[which(results3$significant==TRUE)]
meaning3=t(data.frame(data.frame(t(sapply(as.character(significant_ICD_3), FUN=ukb_icd_code_meaning)))$meaning))
ICD10_3=as.character(meaning3[,1])
final3=data.frame(ICD10_3,significant_pval)
final3 <- final3[order(significant_pval),]
final3

saveRDS(final3 , "/rds/general/project/hda_students_data/live/Group7/General/Final/table_3.rds")
```

```{r}
dim(final3)
```
```{r}
three_year<-ggplot(
  results3, aes(as.numeric(row.names(results3)), y=-log10(pval), color=ICD10))+
  geom_point()+
  geom_hline(yintercept= -log10(0.05/nrow(results3)), linetype="dashed", color = "red")+
  geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(results3)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+
  ggtitle("Time= 3 years prior to diagnosis of skin cancer") +
  xlab("ICD10 codes")+
  ylim(0,50)+
  theme(axis.text.x = element_blank())+
  theme(aspect.ratio = 1)+
   theme(legend.box = 'horizontal')
#+theme(legend.position = "none")

three_year
```
```{r}

ggsave(filename="threeyear.pdf", 
       plot = three_year, 
       device = cairo_pdf, 
       width = 210, 
       height = 297, 
       units = "mm")
```


# 1 year prior to dignosis and above
```{r}
results1$code=rownames(results1)
results1= merge(results1, comorbidities_categories, by='code')
results1<-results1%>%
  drop_na()
results1$ICD10=substr(results1$code, 1, 1)
```

```{r}
results1$logp=-log10(results1$pval)
bonf1=-log10(0.05/nrow(results1))
results1$significant=results1$logp>bonf1
significant_ICD_1=results1$code[which(results1$significant==TRUE)]
significant_pval=results1$pval[which(results1$significant==TRUE)]
meaning1=t(data.frame(data.frame(t(sapply(as.character(significant_ICD_1), FUN=ukb_icd_code_meaning)))$meaning))
ICD10_1=as.character(meaning1[,1])
final1=data.frame(ICD10_1,significant_pval)
final1 <- final1[order(significant_pval),]
final1

saveRDS(final1 , "/rds/general/project/hda_students_data/live/Group7/General/Final/table_1.rds")
```
```{r}
dim(final1)
```

```{r}
one_year<-ggplot(
  results1, aes(as.numeric(row.names(results1)), y=-log10(pval), color=ICD10))+
  geom_point()+
  geom_hline(yintercept= -log10(0.05/nrow(results1)), linetype="dashed", color = "red")+
  geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(results1)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+
  ggtitle("Time= 1 year prior to diagnosis of skin cancer") +
  xlab("ICD10 codes")+
  ylim(0,50)+
  theme(axis.text.x = element_blank())+
  theme(aspect.ratio = 1)+
   theme(legend.box = 'horizontal')
#+theme(legend.position = "none")

one_year
```

```{r}

ggsave(filename="oneyear.pdf", 
       plot = one_year, 
       device = cairo_pdf, 
       width = 210, 
       height = 297, 
       units = "mm")
```


# Right before diagnosis and above 
```{r}
results0$code=rownames(results0)
results0= merge(results0, comorbidities_categories, by='code')
results0<-results0%>%
  drop_na()
results0$ICD10=substr(results0$code, 1, 1)
```

```{r}
results0$logp=-log10(results0$pval)
bonf0=-log10(0.05/nrow(results0))
results0$significant=results0$logp>bonf0
significant_ICD_0=results0$code[which(results0$significant==TRUE)]
significant_pval=results0$pval[which(results0$significant==TRUE)]
meaning0=t(data.frame(data.frame(t(sapply(as.character(significant_ICD_0), FUN=ukb_icd_code_meaning)))$meaning))
ICD10_0=as.character(meaning0[,1])
final0=data.frame(ICD10_0,significant_pval)
final0 <- final0[order(significant_pval),]
final0

saveRDS(final0 , "/rds/general/project/hda_students_data/live/Group7/General/Final/table_0.rds")
```

```{r}
dim(final0)
```


```{r}
zero_year<-ggplot(
  results0, aes(as.numeric(row.names(results0)), y=-log10(pval), color=ICD10))+
  geom_point()+
  geom_hline(yintercept= -log10(0.05/nrow(results0)), linetype="dashed", color = "red")+
  geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(results0)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+
  ggtitle("Time= 0 year prior to diagnosis of skin cancer") +
  xlab("ICD10 codes")+
  ylim(0,50)+
  theme(axis.text.x = element_blank())+
  theme(aspect.ratio = 1)+
   theme(legend.box = 'horizontal')

zero_year
```

```{r}

ggsave(filename="zeroyear.pdf", 
       plot = zero_year, 
       device = cairo_pdf, 
       width = 210, 
       height = 297, 
       units = "mm")
```


# Co-occurent diagnosis

```{r}
results_coc$code=rownames(results_coc)
results_coc= merge(results_coc, comorbidities_categories, by='code')
results_coc<-results_coc%>%
  drop_na()
results_coc$ICD10=substr(results_coc$code, 1, 1)
```


```{r}
results_coc$logp=-log10(results_coc$pval)
bonfco=-log10(0.05/nrow(results_coc))
results_coc$significant=results_coc$logp>bonfco
significant_ICD_co=results_coc$code[which(results_coc$significant==TRUE)]
significant_pval=results_coc$pval[which(results_coc$significant==TRUE)]
meaning0=t(data.frame(data.frame(t(sapply(as.character(significant_ICD_co), FUN=ukb_icd_code_meaning)))$meaning))
ICD10_0=as.character(meaning0[,1])
finalco=data.frame(ICD10_co,significant_pval)
finalco <- finalco[order(significant_pval),]
finalco

saveRDS(finalc , "/rds/general/project/hda_students_data/live/Group7/General/Final/table_co.rds")
```

```{r}
dim(finalco)
```


```{r}
coc<-ggplot(
  results_coc, aes(as.numeric(row.names(results_coc)), y=-log10(pval), color=ICD10))+
  geom_point()+
  geom_hline(yintercept= -log10(0.05/nrow(results_coc)), linetype="dashed", color = "red")+
  geom_text(aes(label=ifelse(-log10(pval)>-log10(0.05/nrow(results_coc)),as.character(code),''),hjust=0,vjust=0),show.legend=FALSE)+
  ggtitle("Time= 0 year prior to diagnosis of skin cancer") +
  xlab("ICD10 codes")+
  ylim(0,50)+
  theme(axis.text.x = element_blank())+
  theme(aspect.ratio = 1)+
  theme(legend.box = 'horizontal')

coc
```

