---
title: "Descriptive analysis"
output: html_document
---

```{r}
library(ggcorrplot)
library(dplyr)

```


Loading cases and controls
```{r}
controls= readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/controls_eid.rds")
cases=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/cases_eid.rds")
```

```{r}
length(controls)
```

```{r}
length(cases)
```

Loading case and controls without dates datasets
```{r}

no_dates_controls= readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/controls_no_dates.rds")
no_dates_cases=readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/cases_no_dates.rds")

```




Loading comorbidity dataset with respective comorbidity percentage
```{r}
comor_df <- readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/disease_counts.rds")
```


Cases and controls presence in comorbidity dataset 
```{r}
cases_not_hes= !(cases %in% rownames(no_dates_cases))
controls_not_hes= !(controls %in% rownames(no_dates_controls))

# finding cases that are not in comorbidity dataset (3 ended up being filtered because they had no date of cancer diag)
cases_no_comor= cases[which(cases_not_hes==TRUE)]
cases_2= cases[which(cases_not_hes==FALSE)]

# finding controls that have/have no comorbidities 
controls_no_comor_eid= controls[which(controls_not_hes==TRUE)]
controls_2= controls[which(controls_not_hes==FALSE)]
```


On average, how many comorbidities have cases and controls?
```{r}
no_dates_controls['sumrow',]<-colSums(no_dates_controls)
no_dates_controls[,'sumcol']<-rowSums(no_dates_controls)

no_dates_cases['sumrow',]<-colSums(no_dates_cases)
no_dates_cases[,'sumcol']<-rowSums(no_dates_cases)

case_row<-no_dates_cases[,'sumcol']
con_row<-no_dates_controls[,'sumcol']

df2=data.frame(eid=cases_2,cases=case_row[1:25127])
names(df2)=c('eids', 'cases')

avg_count_cases= sum(df2$cases)/nrow(df2)


df3=data.frame(eid=controls_2,controls=con_row[1:22271])
names(df3)=c('eids', 'controls')

avg_count_controls= sum(df3$controls)/nrow(df3)

print(avg_count_cases)
print(avg_count_controls)
```

# Analysis of comorbidities in our dataset

```{r}

# ordering the prevelance of comorbidities
cases_ana <- comor_df[order(-comor_df$perc_cases),]
  
skincancer =c(c(paste0("C", 430:449)))

# Removing skin cancer from disease list
skin_filter<-cases_ana[!(rownames(cases_ana) %in% skincancer),]
skin_filter <- skin_filter[order(-skin_filter$perc_cases),]

```


Selecting diseases that are present in more than 5 per cent of cases or controls
```{r}
skin_filter_cases<-skin_filter[which (skin_filter$perc_cases> 0.05),]
skin_filter_controls<-skin_filter[which (skin_filter$perc_controls> 0.05),]
skin_filter_controls<- skin_filter_controls[order(-skin_filter_controls$perc_controls),]

#keeping record of skin cancer diagnosis only
only_skin<-cases_ana[(rownames(cases_ana) %in% skincancer),]

#finding the disease meaning 
only_skin$meaning10_n<-data.frame(t(sapply(as.character(rownames(only_skin)), FUN=ukb_icd_code_meaning)))$meaning

```

#### barplot cases types of melanoma/skin cancer
```{r}
plot1<-ggplot(data=only_skin, aes(x=rownames(only_skin), y=as.numeric(cases))) +
  geom_bar(stat="identity", fill='skyblue') +
  labs(title = "Repartition of cases in skin cancer and melanoma types") + ylab("Number of cases") + xlab("Disease type")+
  coord_flip()

plot1
```
What do these code mean?
```{r}
labels1=unlist(only_skin['meaning10_n'])
print(labels1)
```
# Cases

```{r}
skin_filter_cases$meaning10_n<-data.frame(t(sapply(as.character(rownames(skin_filter_cases)), FUN=ukb_icd_code_meaning)))$meaning
skin_filter_cases= skin_filter_cases[3:56,]


plot2<-ggplot(data=skin_filter_cases, aes(x=rownames(skin_filter_cases), y=as.numeric(cases))) +
  geom_bar(stat="identity", fill='skyblue') +
  labs(title = "Top comorbidities in cases") + ylab("Number of cases") + xlab("Disease type")+
  coord_flip()

plot2

```
What do these labels mean?

```{r}

labels2=unlist(skin_filter_cases['meaning10_n'])
print(labels2)
```
# Controls

```{r}
skin_filter_controls$meaning10_n<-data.frame(t(sapply(as.character(rownames(skin_filter_controls)), FUN=ukb_icd_code_meaning)))$meaning

skin_filter_controls= skin_filter_controls[3:53,]

plot3<-ggplot(data=skin_filter_controls, aes(x=rownames(skin_filter_controls), y=as.numeric(controls))) +
  geom_bar(stat="identity", fill='skyblue') +
  labs(title = "Top comorbidities in controls") + ylab("Number of cases") + xlab("Disease type")+
  coord_flip()

plot3
```

What do these labels mean?

```{r}

labels3=unlist(skin_filter_controls['meaning10_n'])
labels3
```
# Heatmaps 

```{r}

#Carolina to insert code 

```


# Analysis of diagnosis dates 


Loading dates of diagnosis datasets with comorbidities for cases and controls

```{r}
comor_cases <- data.frame(readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/diag_date_recode_ca.rds"))
comor_controls <- data.frame(readRDS("/rds/general/project/hda_students_data/live/Group7/General/Final/diag_date_recode_co.rds"))
```

Filtering dates dataframe based on top prevalences 

```{r}

#most common disease in cases ICD10 codes
icd_cases=rownames(skin_filter_cases)

#filtering dates of diagnosis dataset based on top diseases 
diag_cases<- comor_cases %>%
  select(contains(icd_cases))%>%
  filter(I10!=0)

#editing colnames so it only contains ICD code
colnames(diag_cases) = gsub(pattern = "_date", replacement = "", x = colnames(diag_cases))

table1=data.frame(prop.table(table(diag_cases$I10)))
df_1=matrix(0, ncol = ncol(diag_cases), nrow = 9)
df_1=data.frame(df_1)
rownames(df_1)= table1$Var1
colnames(df_1)=colnames(diag_cases)

for (i in 1:ncol(diag_cases)){
  df_1[,i]=data.frame(prop.table(table(factor(diag_cases[,i], levels=rownames(df_1)))))$Freq
}

```



```{r}
df_1
```
# Same analysis on controls

```{r}

#most common disease in cases ICD10 codes
icd_controls=rownames(skin_filter_controls)

#filtering dates of diagnosis dataset based on top diseases 
diag_controls<- comor_controls %>%
  select(contains(icd_controls))

#editing colnames so it only contains ICD code
colnames(diag_controls) = gsub(pattern = "_date", replacement = "", x = colnames(diag_controls))

table2=data.frame(prop.table(table(diag_controls[,1])))
df_2=matrix(0, ncol = ncol(diag_controls), nrow = 8)
df_2=data.frame(df_2)
rownames(df_2)= table2$Var1
colnames(df_2)=colnames(diag_controls)

for (i in 1:ncol(diag_controls)){
  df_2[,i]=data.frame(prop.table(table(factor(diag_controls[,i], levels=rownames(df_2)))))$Freq
}


```
problem to solve: n=one category of 0 -- need to solve vector using no_skin.R



```{r}
df_2
```

